Improving content discovery through combining linked data and data mining techniques

### 2015-11-10 report/report.tex
Changed title.
Written some background.

\title{Proposal for improving content discovery through combining linked data
and data mining techniques}

@@

\usepackage[autosize]{dot2texi}
\usepackage[Bjarne]{fncychap}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows}

@@

\maketitle

@@

There is some existing work in the literature around data mining and the
semantic web.

@@

\begin{figure}[h]
  \begin{center}
    \begin{dot2tex}[dot,pgf,options=-t math,scale=0.8]
      digraph g {
        rankdir=LR;

        node [shape=circle,margin="0,0"]
        edge [len=2];

@@

\begin{figure}[h]
  \begin{center}
    \begin{dot2tex}[dot,pgf,options=-t math,scale=0.8]
      digraph g {
        rankdir=LR;
        nodesep=20;

@@

\begin{figure}[h]
  \begin{center}
    \begin{dot2tex}[dot,pgf,options=-t math,scale=0.8]
      digraph g {
        rankdir=LR;

@@

>>>>

\title{Improving content discovery through combining linked data and data mining techniques}

@@

\usepackage{nag}
\usepackage{rotating}
\usepackage[autosize]{dot2texi}
\usepackage[Bjarne]{fncychap}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,snakes}
\usepackage{listings}
\usepackage{mathtools}

@@

\maketitle
\tableofcontents

@@

Data mining activities such as machine learning rely on structuring data as
\emph{feature sets}\cite{bishop2006pattern} -- a set or vector of properties or
attributes that describe a single entity. The process of \emph{feature extraction}
generates such feature sets from raw data and is a necessary early phase for
many machine learning activities.

The RDF\footnote{http://www.w3.org/TR/PR-rdf-syntax/} graph is a powerful model
for metadata based on representing knowledge as a set of subject-predicate-object
\emph{triples}. The query language, SPARQL, gives us a way to query the RDF
graph structure using a declarative pattern and return a set of all variable
bindings that satisfy that pattern.

For example, the following SPARQL:

\begin{lstlisting}
  PREFIX foaf: <http://xmlns.com/foaf/0.1/>
  SELECT ?name ?email
  WHERE {
    ?person a foaf:Person.
    ?person foaf:name ?name.
    ?person foaf:mbox ?email.
  }
\end{lstlisting}

\noindent queries an RDF graph that contains contact information and returns the
names and email address of all ``Person'' entities therein. This query could be
expressed as function $f: G \rightarrow S$ from the set of all graphs, $G$, to
set, $S$, of values requested in the SELECT statement such that:

\begin{displaymath}
  S = \{(n, e) | \exists p . (p, type, Person) \in G\ \land (p, name, n) \in G \land (p, mbox, e) \in G\}
\end{displaymath}

This could be generalised:

\begin{displaymath}
  S = \{(a_1, a_2, ..., a_n) | f(G) \}
\end{displaymath}

\noindent and in the case where all $a_k \in S$ are literal (e.g. string or
numeric) values, we can thus consider a given SPARQL query to be specific function
capable of feature extraction from any RDF graph into sets of categorical or
numeric features.

Notably, Kiefer, Bernstein and Locher\cite{kiefer2008adding} proposed a novel
approach called SPARQL-ML -- an extension to the
SPARQL\cite{segaran2009programming} query language with new keywords to
facilitate both generating and applying models. This means that the system capable
of parsing and running standard queries must also run machine learning
algorithms. Their work involved developing an extension to the SPARQL query
engine for \emph{Apache Jena}\footnote{https://jena.apache.org/} that integrates
with systems such as \emph{Weka}\footnote{http://www.cs.waikato.ac.nz/ml/weka/}.
A more suitable software application for enterprise use might focus solely on
converting RDF graphs into a neutral data structure that can plug into arbitrary
data mining algorithms.

@@

\begin{figure}[h]
  \begin{center}
    \begin{dot2tex}[dot,options=-t math,autosize,pgf]
      digraph g {
        rankdir=LR;

        node [shape=circle,margin="0,0"];
        edge [len=2];

@@

\begin{figure}[h]
  \begin{center}
    \begin{dot2tex}[dot,options=-t math,autosize,pgf]
      digraph g {
        rankdir=LR;

@@

\begin{sidewaysfigure}[h]
  \begin{center}
    \begin{dot2tex}[dot,options=-t math,autosize,pgf]
      digraph g {
        rankdir=LR;

        node [shape=circle,margin="0,0"];

        dummy [shape=none,label=""];

        URI [label="URI"];
        RDF1 [label="RDF_1"];
        RDF2 [label="RDF_2"];
        RDFp [label="RDF'"];

        URI -> RDFa [label="fetch"];
        RDFa -> RDF [label="parse"];
        RDF -> RDF1 [label="DBPedia"];
        RDF -> RDF2 [label="DBPedia"];
        RDF -> RDFp [label="\cup"];
        RDF1 -> RDFp [label="\cup"];
        RDF2 -> RDFp [label="\cup"];
        RDFp -> Features [label="generate"];
      }
    \end{dot2tex}
  \end{center}
  \caption{Semantic web content miner with DBPedia enrichment}
\end{sidewaysfigure}

\begin{figure}[h]
  \begin{center}
    \begin{dot2tex}[dot,options=-t math,autosize,pgf]
      digraph g {
        rankdir=LR;

@@

\bibliographystyle{plain}
\bibliography{references}
